An Adolescent Mini-Neptune in the Kepler Field

### 2021-06-15 paper/ms.tex
update ms flares

Kepler\,1627Ab therefore provides an interesting extremum in the ages of the
Kepler planets, and opens multiple avenues for detailed
study.  Ground-based observations of its Doppler transit ($\Delta
v_{\rm RM} \approx 20$\ms) with HIRES or the upcoming KPF spectrograph
could yield a measurement of its orbital obliquity (CITEP,CITEP).
Similarly, transit spectroscopy aimed at detecting either hydrogen or
helium absorption could yield insight into the planet's atmosphere,
which is expected to be outflowing at a rate of {\bf
X.X}\,$M_\oplus\,{\rm yr}^{-1}$ (CITE).  A more challenging
measurement would be a measurement of the planet's mass, through
either detailed timing analyses of the Kepler data, or through a
high-cadence multi-color radial velocity campaign.  The mass would
yield constraints on both the planet's composition (``what is the
extent of the initial gaseous envelope for a typical mini-Neptune?''),
as well as its initial entropy (``hot'' or ``cold''? CITE).

The precision of the Kepler data itself may enable detailed analyses
without even needing to acquire more data.
These could include studies transit timing variations, distortion of
the transit signal due to the star's gravity darkening, 
and interactions between the star and planet exhibited by patterns in
the star's flaring.
On the latter point,
Appendix REFERENCE highlights an odd coincidence: 
in the 97.7 days of available 1-minute cadence Kepler observations
(``short-cadence''), 
24 flares were observed to exceed $0.5\%$ in relative flux.
Cumulatively, these flares spanned a total of 0.27 days.
The brightest flare, highlighted in FIGURE CITE, had a flare occur
after it to within 0.2\% of an orbital period.
This seems {\it a priori} unlikely; in fact, 4 of the 24 flare events
have candidate ``successors'' (events that follow to within 2\% of the
planet's orbital period).
Further analysis would be needed to test the hypothesis of whether or
not the flare time
distribution deviated from a Poisson process...
[like: compare Poisson model versus Poisson+periodic mixture?]

FIXME
(A simpler path to the same result might be to determine whether the
$\approx80\,{\rm ppm}$ residual apparent in our transit fit is caused
by gravity darkening.)

Ultimately, our age measurement for Kepler\,6127Ab was enabled by
identifying its connection to the \cn\ through the Gaia data.  The
Supplementary Table~\ref{tab:v05} enables similar crossmatches to
already known planets, as well as forthcoming planets from the TESS
mission (CITE RickerJATIS2014).  An alternative approach also
likely to bear fruit is to identify kinematic associations
around exoplanet host stars using the Gaia data, and to then verify
these associations with ancilliary rotation periods and spectroscopy
(CITEP Tofflemire).

@@

>>>>

Kepler\,1627Ab therefore provides an interesting extremum in the ages
of the Kepler planets, and opens multiple avenues for study from the
ground.  Observations of its Doppler transit ($\Delta v_{\rm RM}
\approx 20$\ms) with HIRES or the upcoming KPF spectrograph could
yield a measurement of its orbital obliquity (CITEP,CITEP).
Similarly, transit spectroscopy aimed at detecting either hydrogen or
helium absorption could yield insight into the planet's atmosphere,
which is expected to be outflowing at a rate of {\bf
X.X}\,$M_\oplus\,{\rm yr}^{-1}$ (CITE).  A more challenging
measurement would be a measurement of the planet's mass, through
either detailed timing analyses of the Kepler data, or through a
high-cadence multi-color radial velocity campaign.  The mass would
yield constraints on both the planet's composition (``what is the
extent of the initial gaseous envelope for a typical mini-Neptune?''),
as well as its initial entropy (``hot'' or ``cold''? CITE).

The precision of the Kepler data itself may enable more detailed
analyses without even needing to acquire more data.  These could
include studies transit timing variations, distortion of the transit
signal due to the star's gravity darkening, and interactions between
the star and planet exhibited by patterns in the star's flaring.
The gravity darkening might be interesting as an explanation for the
$\approx80\,{\rm ppm}$ residual apparent in our transit fit.
On the point of planet-induced flaring, Appendix REFERENCE
highlights an odd timing coincidence that may be worthy of further
examination.

Ultimately though, our age measurement for Kepler\,6127Ab was enabled by
identifying its connection to the \cn\ through the Gaia data.  The
Supplementary Table~\ref{tab:v05} enables similar crossmatches to
already known planets, as well as forthcoming planets from the TESS
mission (CITE RickerJATIS2014).  An alternative approach also
likely to bear fruit is to identify kinematic associations
around exoplanet host stars using the Gaia data, and to then verify
these associations with ancilliary rotation periods and spectroscopy
(CITEP Tofflemire).

@@

\section{Flare Analysis}
\label{app:flare}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f7.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Flares in Kepler\,1627}.  
    {\it Top:}
    The full short-cadence Kepler dataset, acquired at 1-minute
    sampling (black points) is shown with a stellar variability model
    (blue line).
    {\it Middle:}
    Residual after subtracting the stellar variability model.  Flares
    appear as spikes.
    {\it Bottom:}
    Zooms of the brightest, and third-brightest flares.  A timing
    coincidence -- that both flares has ``successors'' approximately
    one orbital period after the initial event -- is emphasized.
		\label{fig:flarezoom}
	}
\end{figure*}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=1.0\textwidth]{f8.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Phase-folded flares in Kepler\,1627}.  
    {\it Top:}
    As in the middle of Figure~\ref{fig:flarezoom}, phase-folded at
    the planet's ephemeris.
    {\it Bottom:}
    Fitted flare amplitudes and orbital phases.
    Colors indiate relative time, from the beginning of the
    98-day short-cadence Q15 dataset (dark blue) to the end (light
    yellow).
    Lower amplitude flares likely exist in the data, and were not
    examined.
		\label{fig:flarephase}
	}
\end{figure*}

The 1-minute cadence Kepler observations span 98 days, and cover 24
flares exceeding $0.5\%$ in relative flux.  These 24 flares spanned a
total of 6.5 hours ($\sim$15 minutes per flare).  The coincidence is
that despite the low flare duty cycle, one orbital period after the
the brightest flare, a second flare followed.  This and a similar
event are shown in Figure~\ref{fig:flarezoom}.  The timing error is
good to a $\approx0.2\%$ difference from the orbital period, which
seems {\it a priori} unlikely.  If we consider flares falling within
2\% of the planet's orbital period after a previous flare, then 4 of
the 24 flare events have candidate ``successors''.

A brief note on how we cleaned the light curve, and identified the
flares.
For cleaning, we performed the following iterative detrending procedure.
\begin{itemize}
  \item Step 1: Build a two-term mixed SHOTerm GP model with
    quasi-periodic kernels at Prot and 0.5$\times$Prot. Fit the model
    to the time, flux, and flux uncertainty.
  \item Step 2: Select points more than twice the median absolute
    deviation from the residual, and exclude them from the light
    curve.  Repeat Step 1.
  \item Step 3: On the residual from Step 2, identify all flares,
    requiring them to be at least 20 cadences apart, at least 7 median
    absolute deviations above the median baseline, and lasting at least
    2 cadences in duration.  Build the mask spanning these times, from
    5 minutes before each flare begins to 2.5 minutes after the final
    flare cadence.  Repeat Step 1 a final time.
\end{itemize}

The flares were identified and fitted using ALTAIPONY (CITE CITE).
The fitted model is that of CITE XXX, which parametrizes the flare
with a start time, a lag time, and an amplitude (CHECK).
Figure~\ref{fig:flarephase} shows the resulting flares, amplitudes,
and phases.

We considered two hypotheses: {\it i)} the flare arrival times are
distributed as a Poisson process, and {\it ii)} they are better
explained as a Poisson process with a periodic mixture component.

TODO: explore!  Fit the flare inter-arrival times...
[even the 2-separated ones?]
